---
title: "01_setup"
output: html_document
date: "2025-05-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/2298350G/Documents/My Documents/Moths/pre-PhD/E_silaceata/voltinism')
library(dplyr)
library(lubridate)
library(tidyr)
library(ggplot2)
library(viridis)
library(mixtools)
```




# read in data


## data frame containing all moth catches in Rowardennan Rothamsted light trap 1968-2024


```{r}
daily <- read.csv("data/Rowardennan all daily data.csv", header = TRUE, sep = ",")

### subset data to just Ecliptopera silaceata
Esil <- filter(daily, binomial == "Ecliptopera silaceata")


#make binomial (species name) a factor
Esil$binomial <- as.factor(Esil$binomial)
#convert date from character to date format
Esil$CalDate <- dmy(Esil$CalDate)
#make year variable
Esil$Year<-year(Esil$CalDate)
#make julian date variable and convert to numeric
Esil$julian <- strftime(Esil$CalDate, format = "%j")
Esil$julian<-as.numeric(Esil$julian)
```





# assigning number of flight periods to each year



## initial visualisation with heat map



### format data for heat map

```{r}
#make copy of data frame to edit
Esil2 <- Esil
##fill in blank dates in data frame and insert zeroes for absences
Esil2<-  complete(Esil2,CalDate = seq(as.Date("1968-01-01"), as.Date("2024-12-31"), "day"), fill = list(DailyCount = 0))

#add in julian day to new dates with zero abundance
Esil2$julian <- strftime(Esil2$CalDate, format = "%j")
Esil2$julian<-as.numeric(Esil2$julian)

#filter between day of year 100 and 290 for visualisation - all observations fall between these dates
Esil2<-filter(Esil2,julian>100,julian<290)

#update year variable
Esil2$Year<-year(Esil2$CalDate)

#convert back to data frame from tibble
Esil2<-as.data.frame(Esil2)

##log counts for visualisation
Esil2$logcount <- log(Esil2$DailyCount + 1)
```



### make and export heat map

```{r}
#plot heat map
heatmap<- ggplot(Esil2, aes(julian, Year, fill= logcount)) + 
  geom_tile()+
  scale_fill_viridis(discrete=FALSE, name = "Log Abundance")+
  theme_bw()+
  xlab("Day of Year")

heatmap

#export heat map
# ggsave(filename = "outputs/figures/assigning_n_flight_periods/E_silaceata_heat_map.png",
#        plot = heatmap,
#        device = "png",
#        width = 18,
#        height = 12,
#        units = "cm",
#        dpi = 300,
#        bg = "white")

```



Heat map suggests two flight periods - test this using Gaussian mixture model.







## Gaussian mixture model


### format data for gaussian mixture model

```{r}
#create vector where each observation is one individual moth and the value of each observation in the vector is the week of year that the individual was captured
 
#make week variable
Esil$week <- week(Esil$CalDate)



#make empty list
lw_all <- list()

#loop through data repeating the week of year of each catch for the corresponding number of times that there were moths captured during that week
for(i in 1:nrow(Esil)){
  lw_all[[i]] <- rep(Esil[i,12], times = Esil[i,8])
}

#convert list to vector
vw_all<-unlist(lw_all)

```







### run gaussian mixture model 


```{r}
#gaussian mixture model assuming equal split between groups, means of week 22 and week 33, and sigma of 1
mmw <- normalmixEM(vw_all, lambda = .5, mu = c(22, 33), sigma = 1)
summary(mmw)
#density plot of the distributions
plot(mmw, which = 2)
 

#make data frame with posterior probability of each data point belonging to each flight period
pdfw <- as.data.frame(mmw$posterior)


#make vector for each year repeated as many times as there were moths caught in that year
ly <- list()

for(i in 1:nrow(Esil)){
  ly[[i]] <- rep(Esil[i,10], times = Esil[i,8])
}

vy<-unlist(ly)


#add year and week to data set of posterior probabilities for each observation belonging to each flight period
pdfw$Year <- vy
pdfw$week <- vw_all


##write this to a csv - probability of each data point belonging to each flight period
#write.csv(pdfw, "outputs/f_p_probs_each_data_point.csv")


#filter to only observations with high probability of belonging to second flight period
pdfwsub <- filter(pdfw, comp.2 > 0.999)#earliest is week 31
pdfwsub2 <- filter(pdfw, comp.2 <= 0.999)#latest is week 30


#get list of years with high probability of two flight periods
pdfwsub$Year <- as.character(pdfwsub$Year)
pdfwsub$Year <- as.factor(pdfwsub$Year)
years_w_fp <- c(levels(pdfwsub$Year))





### make data set with year and number of flight periods

#make data frame containing all years throughout duration of data set
fpdf <- data.frame(Year = c(1968:2024))


##get annual totals to remove years where no individuals were observed - in these years we cannot be sure how many flight periods there were
annualtotals <- Esil %>% group_by(Year) %>% summarise(total = sum(DailyCount))

#remove years in which no individuals were caught
fpdf <- filter(fpdf, Year %in% annualtotals$Year)

#add number of flight periods in each year
fpdf <- mutate(fpdf, Gen = case_when(
  Year %in% years_w_fp ~ 1,
  .default = 0
))



##export data
#write.csv(fpdf, "outputs/flight_periods.csv")

```



### plot to show data points over the two gaussian distributions

```{r}
#test plot of the two gaussian distributions to assess scale
ggplot(data = data.frame(x = c(15, 40)), aes(x)) +
  stat_function(fun = dnorm, n = 1000, args = list(mean = 22.365332, sd = 1.855219  )) + ylab("") +
  stat_function(fun = dnorm, n = 1000, args = list(mean = 33.327729, sd = 1.855219  )) + ylab("") +
  theme_bw()


##rescale year accordingly to fall between 0 and 0.2
yearadj <- seq(from = 0.00357, to = 0.19992, by = 0.00357)
Year <- c(1968:2023)
yearsdf <- as.data.frame(Year)
yearsdf$yearadj <- yearadj


#convert julian day to week
Esil2$weekj <- Esil2$julian / 7
#values <1 correspond to week 1, therefore the split between flight periods at week 30/31 occurs at any number greater than 30.0

#duplicate data set and add year rescaled
Esil3 <- Esil2
Esil3 <- merge(Esil3, yearsdf, by = "Year")
#filter out zero values (added 1 to count before taking log, so zeroes represent actual zero counts rather than log(1))
Esil4 <- filter(Esil3, logcount > 0)

#add variable for flight period of each point - use week 30/31 cutoff supported by gaussian mixture model
Esil4 <- mutate(Esil4, Gen = case_when(
  weekj > 30 ~ 1,
  .default = 0
))
Esil4$Gen <- as.factor(Esil4$Gen)


## plot gaussian distributions and catch dates 

fp_dates_plot <-ggplot(data = data.frame(x = c(15, 40)), aes(x)) +
  stat_function(fun = dnorm, n = 1000, args = list(mean = 22.365332, sd = 1.855219  )) + 
  stat_function(fun = dnorm, n = 1000, args = list(mean = 33.327729, sd = 1.855219  )) + 
  geom_point(data = Esil4, aes(x = weekj, y = yearadj, shape = Gen))+
  scale_shape_manual(guide = 'none',values=c(1, 16))+
  scale_x_continuous(breaks = c(121/7, 152/7, 182/7, 213/7, 244/7, 274/7), labels = c("1st May.", "1st Jun.", "1st Jul.", "1st Aug.", "1st Sept.", "1st Oct."))+
  scale_y_continuous(breaks = c(0.01071, 0.04641, 0.08211, 0.11781, 0.15351, 0.18921), labels = c("1970", "1980", "1990", "2000", "2010", "2020"))+
  xlab("Date")+
  ylab("Year")+
  theme_classic()

fp_dates_plot

# ggsave(filename = "outputs/figures/assigning_n_flight_periods/fp_dates_distr_plot_bw.png",
#        plot = fp_dates_plot,
#        device = "png",
#        width = 18,
#        height = 12,
#        units = "cm",
#        dpi = 300,
#        bg = "white")

```




