---
title: "02_sliding_window_analysis"
output: html_document
date: "2025-05-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/2298350G/Documents/My Documents/Moths/pre-PhD/E_silaceata/voltinism')
library(dplyr)
library(ggplot2)
library(lmtest)
library(RcppRoll)
library(data.table)
library(pscl)
library(lubridate)
library(DHARMa)
```



## read in data

```{r}
fpdf <- read.csv("outputs/flight_periods.csv", header = TRUE, sep = ",")
fpdf <- subset(fpdf, select = -X)

#temperature data only available up to 2023, so remove 2024 flight period data
fpdf <- filter(fpdf, Year != 2024)
```



## does number of flight periods change with year?


```{r}
yearmod1 <- glm(Gen~ Year, data = fpdf, family = "binomial")
summary(yearmod1)#significant effect

#check residuals
hist(resid(yearmod1))
plot(yearmod1)
yearmod1resid <- simulateResiduals(fittedModel = yearmod1, plot = F)
plotQQunif(yearmod1resid) 
plotResiduals(yearmod1resid)

#run null model
yearmodnull <- glm(Gen ~ 1, data = fpdf, family = "binomial")
summary(yearmodnull)

#likelihood ratio test - is year model better than null model
lrtest(yearmod1,yearmodnull)###year model is significantly better than null model
```




### plot number of flight periods against year using model output

```{r}
yeardf1 <- data.frame(Year = rep(1968:2023))
yearfit1 <- predict(yearmod1, newdata = yeardf1, type = "response", se = TRUE)
yearfit1 <- as.data.frame(yearfit1)
yeardf1$preds<-yearfit1$fit
#add variables for confidence intervals - 2 standard errors
yeardf1$minse<-yearfit1$fit-2*(yearfit1$se.fit)
yeardf1$maxse<-yearfit1$fit+2*(yearfit1$se.fit)


yeargenplot <- ggplot(yeardf1,aes(x=Year,y=preds))+
  geom_ribbon(aes(ymin = minse, ymax =  maxse),alpha=0.25) +
  geom_line(linewidth=1.25)+
  geom_point(data = fpdf, aes(x = Year, y = Gen))+
  theme_classic()+
  xlab("Year")+
  ylab("Probability of Two Flight Periods")


# ggsave(filename = "outputs/figures/year_generation_predict_plot.png",
#        plot = yeargenplot,
#        device = "png",
#        width = 18,
#        height = 12,
#        units = "cm",
#        dpi = 300,
#        bg = "white")
```



## read in temperature data


sourced from HAD-UK Met Office database

```{r}
maxtemp<-read.table("data/max_temp_data.csv", header=T, sep="," )
maxtemp$date<-ymd(maxtemp$date)
maxtemp<-subset(maxtemp,select= -X)

mintemp<-read.table("data/min_temp_data.csv", header=T, sep="," )
mintemp$date<-dmy(mintemp$date)
mintemp<-subset(mintemp,select= -X)

#merge temperature data
alltemp<-merge(mintemp,maxtemp,by="date")
alltemp$Year<-year(alltemp$date)
alltemp$julian <- strftime(alltemp$date, format = "%j")
alltemp$julian<-as.numeric(alltemp$julian)
```





## temperature variables - adjust temperature values to start on day 266 of previous year - latest catch date of second flight period is day 265

```{r}
alltempnoleapyear=subset(alltemp,!(Year %in% c(1968, 1972, 1976, 1980, 1984, 1988, 1992, 1996, 2000, 2004, 2008, 2012, 2016,2020)))
alltempleapyear=subset(alltemp,(Year %in% c(1968, 1972, 1976, 1980, 1984, 1988, 1992, 1996, 2000, 2004, 2008, 2012, 2016,2020)))

#filter to after day 265
alltempnoleapyear2nd<-filter(alltempnoleapyear,julian>265)
#subtract 265 from julian day so that day of year 265 is day zero
alltempnoleapyear2nd$newjulian<-alltempnoleapyear2nd$julian-265
#real year value
alltempnoleapyear2nd$realyear<-alltempnoleapyear2nd$Year
#adjusted year (later half of each year (post-day 265) will now make up the first part of the following year)
alltempnoleapyear2nd$Year<-alltempnoleapyear2nd$Year+1

#same for leap years
alltempleapyear2nd<-filter(alltempleapyear,julian>266)
alltempleapyear2nd$newjulian<-alltempleapyear2nd$julian-266
alltempleapyear2nd$realyear<-alltempleapyear2nd$Year
alltempleapyear2nd$Year<-alltempleapyear2nd$Year+1


#early part of year - pre-day265 - add 100 to julian day (difference between 365 and 265) so that the start of each new year is continuous with the later half of the previous year
alltemp1st<-filter(alltemp,julian<=265)
alltemp1st$newjulian<-alltemp1st$julian+100
alltemp1st$realyear<-alltemp1st$Year

alltempnj1<-rbind(alltempnoleapyear2nd,alltempleapyear2nd,alltemp1st)


## subset to only include dates before emergence of 2nd flight period - the driver of the emergence of the second flight period realistically must occur before the second flight period begins - 

#checking the posterior probabilities of belonging to either flight period from the gaussian mixture model (pdf2 data set generated in script 01) shows that the probability of belonging to flight period 2 is only sufficiently high after week 30 - the mid point of week 30 is day of year 213, so subset the dates to those prior to day 213, newjulian 313

alltempnj<-filter(alltempnj1,newjulian<313)
#write.csv(alltempnj, "outputs/temp_data_formatted.csv")
```




## calculate sliding window temperature variables

### function to calculate sliding temperature windows of varying widths starting on each day of the data  - function adapted from https://stackoverflow.com/questions/18448082/r-data-table-with-rollapply/23533485#23533485 

```{r}
windowed.average <- function(input.table,
                             window.width,
                             id.cols = names(input.table)[2],
                             index.col = names(input.table)[3],
                             val.col = names(input.table)[1]) {
  require(RcppRoll)
#average temperature in a window of given width
  avg.with.group <- 
    input.table[,roll_mean(get(val.col), n = window.width, align = "center"),by=c(id.cols)]
 #average date within this window (gives centre of window)
   avg.index <- 
    input.table[,roll_mean(get(index.col), n = window.width, align = "center"),by=c(id.cols)]$V1

  output.table <- data.table(
    Group = avg.with.group,
    Index = avg.index)

  # rename columns to match inputs
  setnames(output.table, old=colnames(output.table),
           new = c(id.cols,val.col,index.col))
  
  
  #add data on date and window width to data table
  #window width variable
  #w <- paste("w", i, sep = "_")
  #create adjusted julian date variable for each window
output.table$w<- paste("w", window.width, sep = "_")
  output.table$julian<-paste(output.table$w,output.table$newjulian,sep="_")
  
    #create separate column for each window
output.table2<-dcast(output.table, Year ~ julian, value.var = val.col, drop=FALSE)

output.table2 <- subset(output.table2, select = -Year)


}

```




### write function to calculate all temperature window averages for a temperature variable and run all GLMs with each temperature window explaining number of flight periods, make table with window, effect size and R2 to find the best window


```{r}

  swrfunc <- function(tempdata){  
    
#create empty list
datalist = list()

#for loop over different window widths
for(i in c(5:60)){
  
  outputdata <- windowed.average(tempdata, window.width = i)
  
  datalist[[i]] <- outputdata
}

#combine outputs for each window width
  tempwindows = do.call(cbind, datalist)
  #add variable for year
  tempdata$Year <- as.factor(tempdata$Year)
  tempwindows$Year <- levels(tempdata$Year)

  tempwindows <- as.data.frame(tempwindows)
  
  
  
  
  #make data frame including windows and generations
  tempwindows$Year <- as.character(tempwindows$Year)
  tempwindows$Year <- as.numeric(tempwindows$Year)
fptemp<-left_join(fpdf,tempwindows,by="Year")

#glms of gen~tmin for each window
#predictor variables
fptemppredictors<-subset(fptemp,select=-c(Year,Gen))
#number of regressions based on number of windows
tempreg<-ncol(fptemppredictors)
# run n regressions
temp_glms <- lapply(1:tempreg, function(x) glm(fptemp$Gen ~ fptemppredictors[,x],family=binomial))

#list of named characters of pseudo R2 values for each glm
r2temp<-lapply(temp_glms,function(x) pR2(x)[5])
#convert to normal list
r2temp2<-lapply(r2temp, function(x) as.numeric(unlist(x)))
#convert list to vector
r2temp2v <- unlist(r2temp2, use.names = FALSE)
#make data frame with window names
namestemp<-colnames(fptemppredictors)
tempswoutput<-as.data.frame(namestemp)
#add pseudo r2 values to data frame with window names
tempswoutput$r2<-r2temp2v


#add effect size
estemp<-lapply(temp_glms,function(x) coef(summary(x))[2,1])
#convert to normal list
estemp2<-lapply(estemp, function(x) as.numeric(unlist(x)))
#convert list to vector
estemp2v <- unlist(estemp2, use.names = FALSE)
#add effect size values to data frame with window names
tempswoutput$es<-estemp2v


#order by descending pseudo R2 value
tempswoutput <- tempswoutput[rev(order(tempswoutput$r2)),]

}



```






## Sliding time window regression analysis for maximum temperature

```{r}
#get rid of unnecessary variables and format data correctly
tmaxnj<-subset(alltempnj,select=c(tmax,Year,newjulian))
#convert to data table
tmaxnj<-as.data.table(tmaxnj)


#run function
tmaxswoutput <- swrfunc(tmaxnj)


##export glm output summary
#write.csv(tmaxswoutput, "outputs/tmax_sliding_window_glms_r2_effect_size.csv")
```


The best model with the highest R2 (0.28) for maximum temperature used the predictive window w_60_279.5 - a window of 60 days in duration centred on day 279.5 (this is actually day 179.5 as these are the adjusted dates to account for starting the year in the previous year)


The 179th day of the year is the 30th day of this window and day 180 is 31st day;

therefore 179-29 = 150, start day of window is 150;

179 + 30 = 209, end day is 209

60 day window ranging from day of year 150-209



### run the best model for maximum temperature and examine output

```{r}
##create data frame with predictor variables
tmaxwindows <- windowed.average(tmaxnj, window.width = 60)

#add variable for year
tmaxnj$Year <- as.factor(tmaxnj$Year)
tmaxwindows$Year <- levels(tmaxnj$Year)

tmaxwindows <- as.data.frame(tmaxwindows)
  
  
  
#make data frame including windows and generations
tmaxwindows$Year <- as.character(tmaxwindows$Year)
tmaxwindows$Year <- as.numeric(tmaxwindows$Year)
fptmax<-left_join(fpdf,tmaxwindows,by="Year")



#run model
modtmax60_279.5<-glm(Gen~w_60_279.5,data=fptmax,family=binomial)
summary(modtmax60_279.5)
```










## run sliding window function for minimum temperature

```{r}
##minimum temperature data
#get rid of unnecessary variables and format data correctly
tminnj<-subset(alltempnj,select=c(tmin,Year,newjulian))
#convert to data table
tminnj<-as.data.table(tminnj)


tminswoutput <- swrfunc(tminnj)


##export glm output summary
#write.csv(tminswoutput, "outputs/tmin_sliding_window_glms_r2_effect_size.csv")
```

The best model for minimum temperature (R2=0.41) uses a window of 49 day duration centred around day 276 (actually day 176).

Day 176 is the 25th day of this window - the window starts at day 152 and ends at day 200



### run best model for minimum temperature and inspect output

```{r}
##create data frame with predictor variables
tminwindows <- windowed.average(tminnj, window.width = 49)

#add variable for year
tminnj$Year <- as.factor(tminnj$Year)
tminwindows$Year <- levels(tminnj$Year)

tminwindows <- as.data.frame(tminwindows)
  
  
  
#make data frame including windows and generations
tminwindows$Year <- as.character(tminwindows$Year)
tminwindows$Year <- as.numeric(tminwindows$Year)
fptmin<-left_join(fpdf,tminwindows,by="Year")



modtmin49_276<-glm(Gen~w_49_276,data=fptmin,family=binomial)
summary(modtmin49_276)
```



## run sliding window function for mean temperature

```{r}
##make mean temperature variable
alltempnj$tmean <- (alltempnj$tmin + alltempnj$tmax) / 2
#get rid of unnecessary variables and format data correctly
tmeannj<-subset(alltempnj,select=c(tmean,Year,newjulian))
#convert to data table
tmeannj<-as.data.table(tmeannj)


tmeanswoutput <- swrfunc(tmeannj)


##export glm output summary
#write.csv(tmeanswoutput, "outputs/tmean_sliding_window_glms_r2_effect_size.csv")
```

The best model for mean temperature uses a 57-day window centred on day 279 (actually day 179).

Day 179 is 29th day of window - there are 28 days prior to this and 28 days after.
Window start day = 179-28 = 151
Window end day = 179 + 28 = 207




## run best model for mean temperature and inspect output

```{r}
##create data frame with predictor variables
tmeanwindows <- windowed.average(tmeannj, window.width = 57)

#add variable for year
tmeannj$Year <- as.factor(tmeannj$Year)
tmeanwindows$Year <- levels(tmeannj$Year)

tmeanwindows <- as.data.frame(tmeanwindows)
  
  
  
#make data frame including windows and generations
tmeanwindows$Year <- as.character(tmeanwindows$Year)
tmeanwindows$Year <- as.numeric(tmeanwindows$Year)
fptmean<-left_join(fpdf,tmeanwindows,by="Year")


modtmean57_279<-glm(Gen~w_57_279,data=fptmean,family=binomial)
summary(modtmean57_279)
```




## best model is the model using minimum temperature




### check residuals of best model

```{r}
modtmin49_276resid <- simulateResiduals(fittedModel = modtmin49_276, plot = F)
plotQQunif(modtmin49_276resid) 
plotResiduals(modtmin49_276resid)
```






## plot output from best model - minimum temperature window 49_276

```{r}
fptplotdf <- data.frame(w_49_276 = rep(seq(min(fptmin$w_49_276),max(fptmin$w_49_276),by = 0.01), length = 5000))
#get predicted values
fptfit <- predict(modtmin49_276, newdata = fptplotdf, type = "response", se = TRUE)
fptfit <- as.data.frame(fptfit)
fptplotdf$preds<-fptfit$fit
#add variables for confidence intervals - 2 standard errors
fptplotdf$minse<-fptfit$fit-2*(fptfit$se.fit)
fptplotdf$maxse<-fptfit$fit+2*(fptfit$se.fit)


tminplotgen <- ggplot(fptplotdf,aes(x=w_49_276,y=preds))+
  geom_line(aes(x=w_49_276,y=preds))+
  geom_ribbon(aes(ymin = minse, ymax =  maxse),alpha=0.25) +
  geom_point(data=fptmin,aes(x=w_49_276,y=Gen))+
  theme_classic()+
  xlab("Mean Daily Minimum Temperature (\u00B0C) Between Days 152-200")+
  ylab("Probability of Two Flight Periods")+
  theme(plot.margin = margin(10, 12, 10, 10))
tminplotgen

# ggsave(filename = "outputs/figures/tmin_generations_plot.png",
#        plot = tminplotgen,
#        device = "png",
#        width = 18,
#        height = 12,
#        units = "cm",
#        dpi = 300,
#        bg = "white")

```






# cross-validation


## following methods from Simmonds et al (2019), Journal of Animal Ecology



## first test how the number of years of data and the timing of the data (old or recent) affects the model estimate and best window

50 years of data - divide into groups getting sequentially smaller or bigger by ten years - 11 groups ranging from 10 to 50 years in length:
1-10; 1-20; 1-30; 1-40; 1-50; 10-20; 10-30; 10-40; 10-50; 20-30; 20-40; 20-50; 30-40; 30-50; 40-50



### make data subsets

```{r}
#make vector of starting years for data subsets
start <- c(rep(fpdf[1,1], times = 5), rep(fpdf[10,1], times = 4), rep(fpdf[20,1], times = 3), rep(fpdf[30,1], times = 2), rep(fpdf[40,1], times = 1))

#vector of end years for data subsets
end <- c(fpdf[10,1], fpdf[20,1], fpdf[30,1], fpdf[40,1], fpdf[50,1], fpdf[20,1], fpdf[30,1], fpdf[40,1], fpdf[50,1], fpdf[30,1], fpdf[40,1], fpdf[50,1], fpdf[40,1], fpdf[50,1], fpdf[50,1])

#vector of durations
durations <- c(10,20,30,40,50,10,20,30,40,10,20,30,10,20,10)

#put start and end dates into a data frame
startend <- data.frame(start = start, end = end)




##make list of temperature data subsets


#format year correctly
tminnj$Year <- as.character(tminnj$Year)
tminnj$Year <- as.numeric(tminnj$Year)


#make empty list
datasetslist = list()

#for each combination of start and end date, filter the temperature data to fall between these years - just use minimum temperature because it best predicted occurrence of a second flight period
for(i in 1:nrow(startend)){
  subdf <- filter(tminnj, Year >= startend[i,1])
  subdf2 <- filter(subdf, Year <= startend[i,2])
  datasetslist[[i]] <- subdf2
}
```



### run sliding window regression on each data subset

```{r}
#write a function to get the best model for a temperature data set, with effect size and standard error 
  subsamplefunc <- function(tempdata){  

  ##run function for multiple window widths

  #create empty list
datalist = list()

#for loop over different window widths
for(i in c(5:60)){
  
  outputdata <- windowed.average(tempdata, window.width = i)
  
  datalist[[i]] <- outputdata
}

#combine outputs for each window width
  tempwindows = do.call(cbind, datalist)
  #add variable for year
  tempdata$Year <- as.factor(tempdata$Year)
  tempwindows$Year <- levels(tempdata$Year)

  tempwindows <- as.data.frame(tempwindows)
  
  
  #make data frame including windows and generations
  tempwindows$Year <- as.character(tempwindows$Year)
  tempwindows$Year <- as.numeric(tempwindows$Year)
fptemp<-left_join(fpdf,tempwindows,by="Year")

#glms of gen~tmin for each window
#predictor variables
fptemppredictors<-subset(fptemp,select=-c(Year,Gen))
#number of regressions based on number of windows
tempnreg<-ncol(fptemppredictors)
# run n regressions
temp_glms <- lapply(1:tempnreg, function(x) glm(fptemp$Gen ~ fptemppredictors[,x],family=binomial))

#list of named characters of pseudo R2 values for each glm
r2temp<-lapply(temp_glms,function(x) pR2(x)[5])
#convert to normal list
r2temp2<-lapply(r2temp, function(x) as.numeric(unlist(x)))
#convert list to vector
r2temp2v <- unlist(r2temp2, use.names = FALSE)
#make data frame with window names
namestemp<-colnames(fptemppredictors)
tempswoutput<-as.data.frame(namestemp)
#add pseudo r2 values to data frame with window names
tempswoutput$r2<-r2temp2v


#add effect size
estemp<-lapply(temp_glms,function(x) coef(summary(x))[2,1])
#convert to normal list
estemp2<-lapply(estemp, function(x) as.numeric(unlist(x)))
#convert list to vector
estemp2v <- unlist(estemp2, use.names = FALSE)
#add effect size values to data frame with window names
tempswoutput$es<-estemp2v

#add standard error
#add effect size
setemp<-lapply(temp_glms,function(x) coef(summary(x))[2,2])
#convert to normal list
setemp2<-lapply(setemp, function(x) as.numeric(unlist(x)))
#convert list to vector
setemp2v <- unlist(setemp2, use.names = FALSE)
#add effect size values to data frame with window names
tempswoutput$se<-setemp2v

#order by descending pseudo R2 value
tempswoutput <- tempswoutput[rev(order(tempswoutput$r2)),]
#subset to just highest R2 value
tempswoutputfinal <- tempswoutput[1,]

}

  
  #empty list
subsamplelist = list()

##for loop to run sliding window function and output collation for all data subsets
for(i in 1:length(datasetslist)){
  subsampleoutput <- subsamplefunc(datasetslist[[i]])
subsamplelist[[i]] <- subsampleoutput 

}

#make data frame of all outputs
allsubsampleslist = do.call(rbind, subsamplelist)


#add variables for start year, end year, window date and window duration
allsubsamples <- as.data.frame(allsubsampleslist)
allsubsamples$start <- start
allsubsamples$end <- end
allsubsamples$duration <- durations


#add in window duration and timing
allsubsamples$windowmid <- as.numeric(sub(".*\\_(.*?)\\_.*", "\\1", allsubsamples$namestemp))
allsubsamples$windowwidth <- as.numeric(sub(".*\\_(.*?)\\.*", "\\1", substring(allsubsamples$namestemp,3,10)))
```



### test how subsample duration and timing affects results

```{r}
# model to test effect of subsample duration and start date on r2
submod1 <- lm(r2 ~ duration + start, data = allsubsamples)
summary(submod1)#significant effect of duration but not start date


# test effect on effect size
submod2 <- lm(es ~ duration + start, data = allsubsamples)
summary(submod2)#no significant effect of duration, marginally significant effect of start date

submod3 <- lm(se ~ duration + start, data = allsubsamples)
summary(submod3)#estimates more variable with shorter duration

#model for effect on window date
submod4 <- lm(windowmid ~ duration + start, data = allsubsamples)
summary(submod4)#no significant effect of either variable

#effect on window duration
submod5 <- lm(windowwidth ~  duration + start, data = allsubsamples)
summary(submod5)#window width is smaller when calculated with more recent data sets 

```





## K-fold cross validation

```{r}
##function to run k fold cross validation
  kfoldfunc <- function(subtempdata, alltempdata){  

    ##run function for multiple window widths
#create empty list
datalist = list()

#for loop over different window widths - creates sliding temperature windows
for(i in c(5:60)){
  
  outputd2 <- windowed.average(subtempdata, window.width = i)
  
  datalist[[i]] <- outputd2 
}

#combine outputs for each window width
  tempwindows = do.call(cbind, datalist)
  #add variable for year
  subtempdata$Year <- as.factor(subtempdata$Year)
  tempwindows$Year <- levels(subtempdata$Year)

  #sliding windows for minimum temperature
  tempwindows <- as.data.frame(tempwindows)
  

##########################################################  
  ###run for loop to get sliding windows for all years
  #create empty list
datalist2 = list()

#for loop over different window widths - creates sliding temperature windows
for(i in c(5:60)){
  
  outputd2 <- windowed.average(alltempdata, window.width = i)
  
  datalist2[[i]] <- outputd2 
}

#combine outputs for each window width
  alltempwindows = do.call(cbind, datalist2)
  #add variable for year
  alltempdata$Year <- as.factor(alltempdata$Year)
  alltempwindows$Year <- levels(alltempdata$Year)

  #sliding windows 
  alltempwindows <- as.data.frame(alltempwindows)
  
  
#############################################################  

  ###go back to subset data  
  
  #add number of flight periods to temperature window data frame
  #format year correctly
  tempwindows$Year <- as.character(tempwindows$Year)
  tempwindows$Year <- as.numeric(tempwindows$Year)
  #add number of flight periods in each year of the temperature data
fptemp<-left_join(tempwindows,fpdf,by="Year")

#glms of gen~tmin for each window
#subset to just the temperature window predictor variables
fptemppredictors<-subset(fptemp,select=-c(Year,Gen))
#number of regressions based on number of windows
tempnreg<-ncol(fptemppredictors)
# run n regressions
temp_glms <- lapply(1:tempnreg, function(x) glm(fptemp$Gen ~ fptemppredictors[,x],family=binomial))

#list of named characters of pseudo R2 values for each glm
r2temp<-lapply(temp_glms,function(x) pR2(x)[5])
#convert to normal list
r2temp2<-lapply(r2temp, function(x) as.numeric(unlist(x)))
#convert list to vector
r2temp2v <- unlist(r2temp2, use.names = FALSE)

#extract best window name and position
maxr2 <- max(r2temp2v)
position <- match(maxr2, r2temp2v)

#extract temperature window for corresponding position
bestwindowname <- colnames(fptemppredictors[position])
bestwindow <- fptemppredictors[,position]
bestwindowdf <- as.data.frame(bestwindow)
bestwindowdf$Year <- tempwindows$Year

#add flight periods for the years in the temperature data set
fpbestwindow<-left_join(bestwindowdf,fpdf,by="Year")


#re run best model with best temperature model
bestmod <- glm(Gen ~ bestwindow, data = fpbestwindow, family = binomial)

#data frame of temperature in best window in years not included in subsample
tempbw <- alltempwindows[,bestwindowname]
tempbw <- as.data.frame(tempbw)
tempbw <- rename(tempbw, bestwindow = tempbw)
tempbw$Year <- alltempwindows$Year

tempbwother <- filter(tempbw, !Year %in% bestwindowdf$Year)

fppreds <- predict(bestmod,newdata=tempbwother,type="response", se.fit = TRUE)
fppreds<-as.data.frame(fppreds)
tempbwother$preds<-fppreds$fit
tempbwother$se <- fppreds$se.fit
##predicted values on the response scale are the probability of 2 flight periods

fpdfreal <- filter(fpdf, Year %in% tempbwother$Year)
fpdfreal$Year <- as.character(fpdfreal$Year)

tempfppred <- left_join(tempbwother, fpdfreal, by = "Year")
tempfppred$predround <- round(tempfppred$preds)

tempfppred$diff <- tempfppred$Gen - tempfppred$predround
return(tempfppred)

}

  
  
  
  ####define data sets of k fold cross validation
#make sure year is numeric
  #50 years of data - 10 data sets - 10 different sets of 5 years can be removed in turn

startk <- c(fpdf[1,1], fpdf[6,1], fpdf[11,1], fpdf[16,1], fpdf[21,1], fpdf[26,1], fpdf[31,1], fpdf[36,1], fpdf[41,1], fpdf[46,1])

endk <- c(fpdf[5,1], fpdf[10,1], fpdf[15,1], fpdf[20,1], fpdf[25,1], fpdf[30,1], fpdf[35,1], fpdf[40,1], fpdf[45,1], fpdf[50,1])

startendk <- data.frame(startk = startk, endk = endk)

#make sure year is a numeric variable
tminnj$Year <- as.character(tminnj$Year)
tminnj$Year <- as.numeric(tminnj$Year)

##make list of temperature data subsets
kfolddatalist = list()

for(i in 1:nrow(startendk)){
  subdf <- filter(tminnj, !Year %in% c(startendk[i,1]:startendk[i,2]))
  kfolddatalist[[i]] <- subdf
}
  


###run accuracy function for all k fold windows  
 #empty list
kfoldlist = list()

##for loop to run sliding window function and output collation for all data subsets
for(i in 1:length(kfolddatalist)){
  kfoldoutput <- kfoldfunc(kfolddatalist[[i]], tminnj)
kfoldlist[[i]] <- kfoldoutput 

}

allkfoldlist = do.call(rbind, kfoldlist)


#remove rows with NAs
allkfoldlist2 <- na.omit(allkfoldlist)

#check for bias in error - positive or negative number would show bias
sum(allkfoldlist2$diff)

#calculate accuracy
##number of matches
allkfoldlist2$difffac <- as.factor(allkfoldlist2$diff)
nums <- allkfoldlist2 %>%
 group_by(difffac)%>%
 summarise(match = n())
  

#accuracy = number of matches divided by total number of comparisons
accuracy = filter(nums, difffac == 0)[,2] / 50
accuracy

```




## find temperature at which probability of second flight period exceeds 0.5


### using best model = modtmin49_276

effect size = 2.7874
intercept = -27.2523

#minimum temperature in critical window which gives 0.5 probability of occurrence of a second generation
0.5 = 2.7874 * x -27.2523
27.7523 = 2.7874x
x=9.956339







## test whether the critical temperature variable changed over the data set

```{r}
#linear model for effect of year on temperature
tymod1 <- lm(w_49_276 ~ Year, data = tminwindows)
summary(tymod1)#significant effect

#check residuals
plot(tymod1)
hist(resid(tymod1))
tymod1resid <- simulateResiduals(fittedModel = tymod1, plot = F)
plotQQunif(tymod1resid)
plotResiduals(tymod1resid)
```


## plot change in temperature over time

```{r}
tempplotdf1 = data.frame(Year=rep(seq(from = 1968, to = 2023,by=0.1),length=1000))
mintemppreds<-predict(tymod1,newdata=tempplotdf1, se.fit=TRUE,type="response")
mintemppreds<-as.data.frame(mintemppreds)
tempplotdf1$preds<-mintemppreds$fit
#add variables for confidence intervals - 2 standard errors
tempplotdf1$sel<-mintemppreds$fit-(2*mintemppreds$se.fit)
tempplotdf1$seu<-mintemppreds$fit+(2*mintemppreds$se.fit)
tempplotdf1$var <- "min"


##data set with temperature, year and number of generations
fptminsub <- subset(fptmin, select = c("Gen", "Year", "w_49_276"))
fptminsub$Gen <- as.factor(fptminsub$Gen)


#plot temperature over time with number of generations

tempgenplot <- ggplot(tempplotdf1,aes(x=Year,y=preds))+
  geom_ribbon(aes(ymin = sel, ymax =  seu),alpha=0.2) +
  geom_line()+
  geom_point(data=fptminsub,aes(x=Year,y=w_49_276, shape = Gen))+
  theme_classic()+
  scale_shape_manual(name = "Number of\nFlight Periods", labels = c("1", "2"), values = c(1, 16))+
  ylab("Temperature in Critical Window")
tempgenplot


# ggsave(filename = "outputs/figures/temp_year_fp_plot_bw.png",
#        plot = tempgenplot,
#        device = "png",
#        width = 18,
#        height = 12,
#        units = "cm",
#        dpi = 300,
#        bg = "white")
```




