---
title: "02_sliding_window_analysis"
output: html_document
date: "2025-05-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/2298350G/Documents/My Documents/Moths/pre-PhD/E_silaceata/voltinism')
library(dplyr)
library(ggplot2)
library(lmtest)
library(RcppRoll)
library(data.table)
library(pscl)
```



## read in data

```{r}
fpdf <- read.csv("outputs/flight_periods.csv", header = TRUE, sep = ",")
fpdf <- subset(fpdf, select = -X)
```



## does number of flight periods change with year?


```{r}
yearmod1 <- glm(Gen~ Year, data = fpdf, family = "binomial")
summary(yearmod1)#significant effect

#check residuals
hist(resid(yearmod1))
plot(yearmod1)

#run null model
yearmodnull <- glm(Gen ~ 1, data = fpdf, family = "binomial")
summary(yearmodnull)

#likelihood ratio test - is year model better than null model
lrtest(yearmod1,yearmodnull)###year model is significantly better than null model
```


## read in temperature data


sourced from HAD-UK Met Office database

```{r}
maxtemp<-read.table("data/max_temp_data.csv", header=T, sep="," )
maxtemp$date<-ymd(maxtemp$date)
maxtemp<-subset(maxtemp,select= -X)

mintemp<-read.table("data/min_temp_data.csv", header=T, sep="," )
mintemp$date<-dmy(mintemp$date)
mintemp<-subset(mintemp,select= -X)

#merge temperature data
alltemp<-merge(mintemp,maxtemp,by="date")
alltemp$Year<-year(alltemp$date)
alltemp$julian <- strftime(alltemp$date, format = "%j")
alltemp$julian<-as.numeric(alltemp$julian)
```





## temperature variables - adjust temperature values to start on day 266 of previous year - latest catch date of second flight period is day 265

```{r}
alltempnoleapyear=subset(alltemp,!(Year %in% c(1968, 1972, 1976, 1980, 1984, 1988, 1992, 1996, 2000, 2004, 2008, 2012, 2016,2020)))
alltempleapyear=subset(alltemp,(Year %in% c(1968, 1972, 1976, 1980, 1984, 1988, 1992, 1996, 2000, 2004, 2008, 2012, 2016,2020)))

#filter to after day 265
alltempnoleapyear2nd<-filter(alltempnoleapyear,julian>265)
#subtract 265 from julian day so that day of year 265 is day zero
alltempnoleapyear2nd$newjulian<-alltempnoleapyear2nd$julian-265
#real year value
alltempnoleapyear2nd$realyear<-alltempnoleapyear2nd$Year
#adjusted year (later half of each year (post-day 265) will now make up the first part of the following year)
alltempnoleapyear2nd$Year<-alltempnoleapyear2nd$Year+1

#same for leap years
alltempleapyear2nd<-filter(alltempleapyear,julian>266)
alltempleapyear2nd$newjulian<-alltempleapyear2nd$julian-266
alltempleapyear2nd$realyear<-alltempleapyear2nd$Year
alltempleapyear2nd$Year<-alltempleapyear2nd$Year+1


#early part of year - pre-day265 - add 100 to julian day (difference between 365 and 265) so that the start of each new year is continuous with the later half of the previous year
alltemp1st<-filter(alltemp,julian<=265)
alltemp1st$newjulian<-alltemp1st$julian+100
alltemp1st$realyear<-alltemp1st$Year

alltempnj1<-rbind(alltempnoleapyear2nd,alltempleapyear2nd,alltemp1st)


## subset to only include dates before emergence of 2nd flight period - the driver of the emergence of the second flight period realistically must occur before the second flight period begins - 

#checking the posterior probabilities of belonging to either flight period from the gaussian mixture model (pdf2 data set generated in script 01) shows that the probability of belonging to flight period 2 is only sufficiently high after week 30 - the mid point of week 30 is day of year 213, so subset the dates to those prior to day 213, newjulian 313

alltempnj<-filter(alltempnj1,newjulian<313)
```




## calculate sliding window temperature variables

### function to calculate sliding temperature windows of varying widths starting on each day of the data  - function adapted from https://stackoverflow.com/questions/18448082/r-data-table-with-rollapply/23533485#23533485 

```{r}
windowed.average <- function(input.table,
                             window.width,
                             id.cols = names(input.table)[2],
                             index.col = names(input.table)[3],
                             val.col = names(input.table)[1]) {
  require(RcppRoll)
#average temperature in a window of given width
  avg.with.group <- 
    input.table[,roll_mean(get(val.col), n = window.width, align = "center"),by=c(id.cols)]
 #average date within this window (gives centre of window)
   avg.index <- 
    input.table[,roll_mean(get(index.col), n = window.width, align = "center"),by=c(id.cols)]$V1

  output.table <- data.table(
    Group = avg.with.group,
    Index = avg.index)

  # rename columns to match inputs
  setnames(output.table, old=colnames(output.table),
           new = c(id.cols,val.col,index.col))
  
  
  #add data on date and window width to data table
  #window width variable
  #w <- paste("w", i, sep = "_")
  #create adjusted julian date variable for each window
output.table$w<- paste("w", window.width, sep = "_")
  output.table$julian<-paste(output.table$w,output.table$newjulian,sep="_")
  
    #create separate column for each window
output.table2<-dcast(output.table, Year ~ julian, value.var = val.col, drop=FALSE)

output.table2 <- subset(output.table2, select = -Year)


}

```




## run function for maximum temperature data


### create sliding windows for all days with window length varying from 5-60 days

```{r}
##maximum temperature data
#get rid of unnecessary variables and format data correctly
tmaxnj<-subset(alltempnj,select=c(tmax,Year,newjulian))
#convert to data table
tmaxnj<-as.data.table(tmaxnj)




##run function for multiple window widths

#create empty list
datalist = list()

#for loop over different window widths
for(i in c(5:60)){
  
  outputd2 <- windowed.average(tmaxnj, window.width = i)
  
  datalist[[i]] <- outputd2 
}

#combine outputs for each window width
  tmaxwindows = do.call(cbind, datalist)
  #add variable for year
  tmaxwindows$Year <- 1968:2024

  tmaxwindows <- as.data.frame(tmaxwindows)
```





## run GLMs on maximum temperature window data

```{r}
#make data frame including windows and generations
fptmax<-left_join(fpdf,tmaxwindows,by="Year")

#glms of gen~tmax for each window
#predictor variables
fptmaxpredictors<-subset(fptmax,select=-c(Year,Gen))
#number of regressions based on number of windows
tmaxnreg<-ncol(fptmaxpredictors)
# run n regressions
tmax_glms <- lapply(1:tmaxnreg, function(x) glm(fptmax$Gen ~ fptmaxpredictors[,x],family=binomial))

#list of named characters of pseudo R2 values for each glm
r2tmax<-lapply(tmax_glms,function(x) pR2(x)[5])
#convert to normal list
r2tmax2<-lapply(r2tmax, function(x) as.numeric(unlist(x)))
#convert list to vector
r2tmax2v <- unlist(r2tmax2, use.names = FALSE)
#make data frame with window names
namestmax<-colnames(fptmaxpredictors)
tmaxswoutput<-as.data.frame(namestmax)
#add pseudo r2 values to data frame with window names
tmaxswoutput$r2<-r2tmax2v


#add effect size
estmax<-lapply(tmax_glms,function(x) coef(summary(x))[2,1])
#convert to normal list
estmax2<-lapply(estmax, function(x) as.numeric(unlist(x)))
#convert list to vector
estmax2v <- unlist(estmax2, use.names = FALSE)
#add effect size values to data frame with window names
tmaxswoutput$es<-estmax2v
#order by descending pseudo R2 value
tmaxswoutput <- tmaxswoutput[rev(order(tmaxswoutput$r2)),]



##export glm output summary
write.csv(tmaxswoutput, "outputs/tmax_sliding_window_glms_r2_effect_size.csv")


##best model - highest R2
#w_5_146
#window length 5 days
#centred on day 146
modtmax5_146<-glm(Gen~w_5_146,data=fptmax,family=binomial)
summary(modtmax5_146)
#day 146 - actually 46 - starts day 44, ends day 48 

```




## make sliding windows for minimum temperature 

### window width varying from 5-60 days

```{r}
##minimum temperature data
#get rid of unnecessary variables and format data correctly
tminnj<-subset(alltempnj,select=c(tmin,Year,newjulian))
#convert to data table
tminnj<-as.data.table(tminnj)




##run function for multiple window widths

#create empty list
datalist = list()

#for loop over different window widths
for(i in c(5:60)){
  
  outputd2 <- windowed.average(tminnj, window.width = i)
  
  datalist[[i]] <- outputd2 
}

#combine outputs for each window width
  tminwindows = do.call(cbind, datalist)
  #add variable for year
  tminwindows$Year <- 1968:2024

  tminwindows <- as.data.frame(tminwindows)
  
  
  
  
  #make data frame including windows and generations
fptmin<-left_join(fpdf,tminwindows,by="Year")

#glms of gen~tmin for each window
#predictor variables
fptminpredictors<-subset(fptmin,select=-c(Year,Gen))
#number of regressions based on number of windows
tminnreg<-ncol(fptminpredictors)
# run n regressions
tmin_glms <- lapply(1:tminnreg, function(x) glm(fptmin$Gen ~ fptminpredictors[,x],family=binomial))

#list of named characters of pseudo R2 values for each glm
r2tmin<-lapply(tmin_glms,function(x) pR2(x)[5])
#convert to normal list
r2tmin2<-lapply(r2tmin, function(x) as.numeric(unlist(x)))
#convert list to vector
r2tmin2v <- unlist(r2tmin2, use.names = FALSE)
#make data frame with window names
namestmin<-colnames(fptminpredictors)
tminswoutput<-as.data.frame(namestmin)
#add pseudo r2 values to data frame with window names
tminswoutput$r2<-r2tmin2v



#add effect size
estmin<-lapply(tmin_glms,function(x) coef(summary(x))[2,1])
#convert to normal list
estmin2<-lapply(estmin, function(x) as.numeric(unlist(x)))
#convert list to vector
estmin2v <- unlist(estmin2, use.names = FALSE)
#add effect size values to data frame with window names
tminswoutput$es<-estmin2v
#order by descending pseudo R2 value
tminswoutput <- tminswoutput[rev(order(tminswoutput$r2)),]



##export glm output summary
write.csv(tminswoutput, "outputs/tmin_sliding_window_glms_r2_effect_size.csv")


##best model
#w_49_276
#window length 49 days
#centred on day 276
modtmin49_276<-glm(Gen~w_49_276,data=fptmin,family=binomial)
summary(modtmin49_276)
#day 276 - actually 176 - day 176 is day 25 - window starts day 152, ends day 200 
```



## mean temperature sliding windows

```{r}
##mean temperature data
alltempnj$tmean <- (alltempnj$tmin + alltempnj$tmax) / 2
#get rid of unnecessary variables and format data correctly
tmeannj<-subset(alltempnj,select=c(tmean,Year,newjulian))
#convert to data table
tmeannj<-as.data.table(tmeannj)




##run function for multiple window widths

#create empty list
datalist = list()

#for loop over different window widths
for(i in c(5:60)){
  
  outputd2 <- windowed.average(tmeannj, window.width = i)
  
  datalist[[i]] <- outputd2 
}

#combine outputs for each window width
  tmeanwindows = do.call(cbind, datalist)
  #add variable for year
  tmeanwindows$Year <- 1968:2024

  tmeanwindows <- as.data.frame(tmeanwindows)
  
  
  
  
  #make data frame including windows and generations
fptmean<-left_join(fpdf,tmeanwindows,by="Year")

#glms of gen~tmean for each window
#predictor variables
fptmeanpredictors<-subset(fptmean,select=-c(Year,Gen))
#number of regressions based on number of windows
tmeannreg<-ncol(fptmeanpredictors)
# run n regressions
tmean_glms <- lapply(1:tmeannreg, function(x) glm(fptmean$Gen ~ fptmeanpredictors[,x],family=binomial))

#list of named characters of pseudo R2 values for each glm
r2tmean<-lapply(tmean_glms,function(x) pR2(x)[5])
#convert to normal list
r2tmean2<-lapply(r2tmean, function(x) as.numeric(unlist(x)))
#convert list to vector
r2tmean2v <- unlist(r2tmean2, use.names = FALSE)
#make data frame with window names
namestmean<-colnames(fptmeanpredictors)
tmeanswoutput<-as.data.frame(namestmean)
#add pseudo r2 values to data frame with window names
tmeanswoutput$r2<-r2tmean2v



#add effect size
estmean<-lapply(tmean_glms,function(x) coef(summary(x))[2,1])
#convert to normal list
estmean2<-lapply(estmean, function(x) as.numeric(unlist(x)))
#convert list to vector
estmean2v <- unlist(estmean2, use.names = FALSE)
#add effect size values to data frame with window names
tmeanswoutput$es<-estmean2v
#order by descending pseudo R2 value
tmeanswoutput <- tmeanswoutput[rev(order(tmeanswoutput$r2)),]


##export glm output summary
write.csv(tmeanswoutput, "outputs/tmean_sliding_window_glms_r2_effect_size.csv")


##best model
#w_22_287.5
#window length 22 days
#centred on day 287.5
modtmean22_287.5<-glm(Gen~w_22_287.5,data=fptmean,family=binomial)
summary(modtmean22_287.5)
#day 287.5 - actually 187.5 - day 187 is day 11 - window starts day 177, ends day 198 
```


