---
title: "03_abundance"
output: html_document
date: "2025-05-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/2298350G/Documents/My Documents/Moths/pre-PhD/E_silaceata/voltinism')
library(dplyr)
library(ggplot2)
library(lmtest)
library(glmmTMB)
library(lme4)
library(DHARMa)
```




## read in data containing one row for each individual E silaceata caught and the probability of it belonging to each flight period and use this to calculate abundance of each flight period in each year

```{r}
probs <- read.csv("outputs/f_p_probs_each_data_point.csv", header = TRUE, sep = ",")
probs$Year <- as.factor(as.character(probs$Year))

#create new variable for flight period assignment
probs$Gen <- as.factor(round(probs$comp.2))

#create new data set with abundance of each flight period in each year
abun <- probs %>%
  group_by(Year, Gen) %>%
  summarise(count = n())

##add in the zero counts for 1972, 1996 and 1997 - these can be treated as actual data points rather than zeroes due to error

#make data set
abunall <- abun
abunzeroes <- data.frame(Year = c("1972", "1996", "1997", "2019"), Gen = c("0", "0", "0", "0"), count = c(0,0,0,0))
#combine
abunall <- rbind(abunall,abunzeroes)

```




## make data frame with abundance of first generation in each year, abundance of previous generation, and whether previous generation was first or second


```{r}
#add rows for all years
Year <- as.factor(1968:2024)
allyearsdf <- as.data.frame(Year)

abun2 <- left_join(allyearsdf, abunall, by = "Year")

#create variable for abundance of previous generation
abun2$Ntminusone <- c(NA,head(abun2$count, -1))

#create variable for phenology of previous generation
abun2$Gentminusone <- as.factor(c(NA,head(as.character(abun2$Gen), -1)))

#subset data to only include first generations in each year
abun1 <- filter(abun2, Gen == 0)

```






```{r}
alltempnj <- read.csv("outputs/temp_data_formatted.csv", sep = ",", header = TRUE)
alltempnj$date <- ymd(alltempnj$date)

#filter to dates between october and june
subtemp <- alltempnj %>%
  filter(newjulian %in% c(9:282)) %>%
  group_by(Year) %>%
  summarise(tmean = mean((tmin+tmax)/2))


#add temperature to abundance data
abun1t <- merge(abun1, subtemp, by = "Year")
#remove NA values
abun1tsub <- na.omit(abun1t)

```



## model abundance of first generation in each year against previous generation abundance and voltinism and average temperature october to june

```{r}
##model selection

#full model with temperature, abundance of previous generation, voltinism of previous generation, and interaction
trmod1test2 <- glmmTMB(count ~ tmean + Ntminusone + Gentminusone + Ntminusone * Gentminusone,data = abun1tsub, family=nbinom1)
summary(trmod1test2)

#model without interaction
trmod2test2 <- glmmTMB(count ~ tmean + Ntminusone + Gentminusone,data = abun1tsub, family=nbinom1)
summary(trmod2test2)

#LRT to test significance of interaction
lrtest(trmod1test2, trmod2test2)#keep simple model

#model without voltinism of previous generation
trmod3test2 <- glmmTMB(count ~ tmean + Ntminusone,data = abun1tsub, family=nbinom1)
summary(trmod3test2)
#LRT for significance of voltinism of previous generation
lrtest(trmod2test2,trmod3test2)#keep simple model

#model without abundance of previous generation
trmod4test2 <- glmmTMB(count ~ tmean ,data = abun1tsub, family=nbinom1)
summary(trmod4test2)
#LRT for significance of abundance of previous generation
lrtest(trmod3test2,trmod4test2)#keep complex model

#model without temperature
trmod5test2 <- glmmTMB(count ~ Ntminusone,data = abun1tsub, family=nbinom1)
summary(trmod5test2)#keep complex model
#LRT for significance of temperature
lrtest(trmod3test2,trmod5test2)#marginally significant - keep complex model


#examine final model
summary(trmod3test2)


##check residuals of final model
trmod3test2resid <- simulateResiduals(fittedModel = trmod3test2, plot = F)
plotQQunif(trmod3test2resid)
plotResiduals(trmod3test2resid)
```



